---
title: "Crop Yield - R Plumber API Wrapper"
author: "Josiah Parry"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
library(dplyr)
```

## Querying an API


We use the `{httr}` library for interacting with HTTP requests from R. 

Generate the API endpoint query URL.

```{r}
library(httr)

# Query API w/out function ------------------------------------------------

query_url <- modify_url(
  url = "https://colorado.rstudio.com/rsc/crop-yield-api/crop-yield",
  # these are the parameters that are sent to the API end point 
  # this will become function arguments 
  query = list(.year = 1999, .product = "maize", .entity = "united states")
)

query_url
```

Send the query to the API using the correct _method_. In this case, `httr::GET()`.  

```{r}
# we want to send the query
res <- GET(query_url)

res
```
We see we've got a status code of 200 which is ideal! Now we need to extract the contents of the API query result. Use `httr::content()` to get the results from the API. We specify the argument `as = "text"` so we can get back the raw JSON that the API sent us. Otherwise, httr will guess and parse the result for us making a list object. I'd rather we parse it ourselves into a data.frame! We are also specifying the text encoding to prevent httr from making an educated guess. Being more specific is better.

```{r}
# parse the query response as text so we get the raw json
res_json <- content(res, as = "text", encoding = "UTF-8")

cat(res_json)
```

Notice that we have JSON now. We can use the fabulous `jsonlite` package to parse a character string which contains JSON. This will read the results into a dataframe. 

```{r}
# read the json into a data frame 
jsonlite::fromJSON(res_json)
```

## Making a wrapper

The above was good, but it's not easily repeatable for multiple queries. To make life easier, we will bundle all of the above code up into a single function. 


```{r}
# Create plumber API wrapper  ---------------------------------------------


get_crop_yields <- function(.year, .product, .entity) {
  
  query_url <- modify_url(
    url = "https://colorado.rstudio.com/rsc/crop-yield-api/crop-yield",
  
      # Function arguments are passed into the `query` argument to fill out the parameters
    # key-value pairs. 
    
    query = list(.year = .year, .product = .product, .entity = .entity)
  )
  
  res <- GET(query_url)
  
  res_json <- content(res, as = "text", encoding = "UTF-8")
  
  jsonlite::fromJSON(res_json)
  
}
```

Use our newly defined function. 

```{r}
get_crop_yields(2012, "beans", "mexico")
```

This function is extremely useful because now we can iterate over multiple values! 

```{r}
to_query <- expand.grid(.year = 2010:2012,
            .product = c("beans", "maize"),
            .entity = "united states")

results <- purrr::pmap_dfr(to_query, get_crop_yields)

results
```

```{r warning=FALSE, message=FALSE}
library(ggplot2)

ggplot(results, aes(year, crop_yield)) +
  geom_col(position = "dodge", size = 0.8) +
  facet_wrap("product", scales = "free")
```

